---
title: "Specimens per species - supplemental analysis"
output:
  pdf_document: default
  html_document: default
---
  
  ```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

```{r}
# Preliminaries

# Load libraries
library(tidyverse)
library(patchwork) # from GitHub
library(png)
library(knitr)
library(broom)
library(here)

#install.packages("remotes")
#remotes::install_github("sckott/rphylopic")
library(rphylopic)

# Read in phylopics
img_mouse <- readPNG(here("img/mouse.png"))
img_tit <- readPNG(here("img/tit.png"))

# Colour blind friendly palette
cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", 
               "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

# Helper functions for plotting
remove_y <- 
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())

remove_x <-   
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())
```

```{r}
# Read in all the specimen data
specimens <- read_csv(here("data/all-specimen-data.csv")) 
```

## How does the proportion of male and female specimens vary as we increase the number of specimens?
We expect large skews in the percentage of male or female specimens when sample size is low, but the ratio of male:female specimens should approach 50:50 as more specimens are added (assuming a 50:50 sex ratio in nature). 

Most species are repesented by only a few specimens (see histograms below), with large skews in % female specimens (in both directions) at low numbers.

```{r, fig.cap = "Distribution of percentage female specimens and log number of specimens per species across birds and mammals."}
source(here("analyses/figure-specimens-all.R"))
hist1 / hist2
```
```{r, fig.cap = "Relationship between the percentage female specimens per species and the number of specimens per species. Hex bins are used rather than points to make the plot clearer."}
all
```

If we fit a model (LM or GLM) to this data we see massive heteroscedasticity in the residuals.
```{r, fig.cap = "Model diagnostic plots for models of % female specimens as a function of log total number of specimens for each species"}
ds_females <-
  specimens %>%
  filter(!is.na(sex)) %>%
  group_by(class) %>%
  add_count(binomial) %>%
  add_count(binomial, sex) %>%
  select(class, binomial, sex, n, nn) %>%
  distinct() %>%
  mutate(percent = round(nn/n*100, 2)) %>%
  filter(sex == "Female")

spec_model <- lm(percent ~ log(n), data = ds_females)

# Diagnostics
par(mfrow = c(2, 2))
plot(spec_model, add.smooth = FALSE)
par(mfrow = c(1, 1))
```

To avoid this variability influencing our results we excluded all species with fewer than 100 specimens from our analyses. Model fit improved substantially (see below).

```{r, fig.cap = "Model diagnostic plots for models of % female specimens as a function of log total number of specimens for each species, for species with > 100 specimens"}
ds100 <- 
  ds_females %>%
  filter(n >= 100)

spec_model2 <- lm(percent ~ log(n), data = ds100)

# Diagnostics
par(mfrow = c(2, 2))
plot(spec_model2, add.smooth = FALSE)
par(mfrow = c(1, 1))
```

There is still a significant positive relationship between the number of specimens and the percentage of female specimens, but the effect size is extremely low (slope = `r round(tidy(summary(spec_model2))$estimate[2], 3)`, se = `r round(tidy(summary(spec_model2))$std.error[2], 3)`, t = `r round(tidy(summary(spec_model2))$statistic[2], 3)`,  df = `r tidy(anova(spec_model2))[2, 2]`, p = `r round(tidy(summary(spec_model2))$p.value[2], 3)`, r2 = `r round(glance(summary(spec_model2))$adj.r.squared, 3)`) so we exclude this variable from future analyses. 